FraCaS: 80 percent complete

### 2019-01-21 paper2018/paper.tex
revision

\usepackage{stmaryrd}
\usepackage{newunicodechar}
\input{PaperTools/latex/newunicodedefs}

@@

@@

The dynamic semantics follows a monadic structure, as pioneered by
\citet{Shan:2002}. There are two kinds of effects carried by the
monad.  The first one comprises a series of state updates and queries,
corresponding to all elements that can be referred to by anaphoric
expressions. These can be usual ones (like NPs), but also less usual
ones (like 2-place verbs, or a quantity --- which we illustrate
below). The other kind of effects is \emph{non-determinism}. This
models the property that linguistic expressions can have several
interpretations. The monadic structure allows to locally express that
a given expression has several meaning; the monadic bind ensures that
all combinations of meanings are considered at the top-level,
combinatorially. This dynamic semantics allows us to model many
phenomena in a precise way:

\paragraph{anaphora} We can handle anaphora % wait so see what JoLLi says here

  % - Dynamic Semantics/Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
%   -> Gender is a property of nouns
\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase.

\paragraph{definites} A naive way to handle definites is as sigma (or
exsistential types). \todo{add example} However, if the semantics
does not feature a dynamic element, then the existential
quantification is introduced locally. This means that the quantifier
can be introduced in the wrong context. Consider the phrase
``everyone pets the dog''. The structure of the interpretation would
be be $∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

Instead, our take is that definites should be treated as anaphoric
expressions. If the referent is not found in the discourse, then it
will be introduced using a sigma type, \emph{at the top-level} of
the expression. Therefore we record all definites without referent,
using a monadic effect. When the complete sentence has been
interpreted, we re-introduced referent using existentials. For the
above example, we obtain the desired interpretation:
be be $∃y. dog(y) ∧ ∀x. person(x) → pet(x,y)$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle section of the FraCaS test suite showed
that handling comparatives is not easy. Our strategy here is to
leverage our dynamic semantics, revealing an anaphoric element of
comparatives. Consider the hypothesis of FraCaS 239: ``ITEL won more
orders than APCOM lost.''  We postulate that this sentence is
equivalent to the following two separate parts: ``APCOM lost zero or
more orders. ITEL won more orders [than some elliptic quantity].''
The elliptic quantity is introduced every time we talk about some
quantity (indexed by a CN, in this case ``orders''), and it can be
referred to by a comparative, later in the discourse. Using this idea,
we can go one level deeper in the interpretation of our example:
``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1 orders.''. We
see here how the quantities are introduced. They are added to the
environment so that, they can be referred to as elliptic quantity
expressions. Finally, ``more'' is intepreted as ``at least <elliptic
quantity>+1''.

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
\todo{reference to this view in the litterature}.  Additionally,
adjectival assumptions as given in FraCaS are interpreted as linear
equations in $Z$. Because solving such equations is decidable, solving
such problems can be made efficient. Indeed, the tactic that Coq
offers for this purpose can solve all such problems in FraCaS,
automatically.

\paragraph{Adverbs}
Another point, of minor theoretical importance but major practical
one, is our handling of adverbial phrases. We interpret adverbs (and
in general all adverbial and prepositional phrases) are VP-modifiers:
$Adv = VP → VP$, where $VP = object → Prop$. However, applying
adverbs to verb-phrases heavily complicates the Coq proofs, because
such phrases can contain quantifiers. Therefore, we instead move the
adverbs, so that they apply to (atomic) verbs only, and can simplify
proofs accordingly. \todo{theoretical justification in the literature}

@@

We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
total of 259 examples. We excluded only one section, 7. The reason for
doing so is that, in our view, it contains too many examples which
require \textit{ad-hoc} treatment, and thus makes little sense to
include without complementing it with a more thorough suite which
captures a more complete landscape of the phenomena that section 7
touches.

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  For this
work, we have amended FraCaS to address a few problems. First, certain
test case are not formed correctly. Those were identified by Bill
MacCartney as ``undef'', and we removed those. Second, a few test
cases occurs twice in the suite, but with two different answers (one
YES and one UNK), with the annotation that those correspond to
different readings. However, elsewhere in the suite, if a problem has
several readings but only one has entailment, it occurs only once and
is marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. Finally, we
have removed case 199 (which appears to be vacuous), and changed the
following cases, which appeared to have been misclassified. We note
that the majority of the mistaken classifications occur in sections 3
and 4, which have not been previously attempted and thus, we propose,
have not been properly scrutinized.

@@

Our system classifies a case as YES if a proof can be constructed from
the premises to the hypothesis, NO if a proof of the negated
hypothesis can be constructed and UNK otherwise. Because we work with
a non-decidable logic, one can not \emph{in general} conclude
decisively that no proof exists. Thus, we consider here that no proof
exists if it cannot be constructed with reasonable effort. In
particular, we test at the minimum that the automatic proof search
built in Coq does not succeed before classifying a problem as UNK.

@@

\providecommand\forcecenter{\multicolumn{1}{c}}
\begin{table}
  \centering
\begin{tabular}{rlrrrrrr}
  & Section      & \# examples & Ours     & FraCoq & MINE & Nut  & Langpro  \\ \hline
1 & Quantifiers  & 75          & .96 (74) & .96    & .77  & .53  & .93 (44) \\
2 & Plurals      & 33          & .82      & .76    & .67  & .52  & .73 (24) \\
3 & Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
4 & Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
5 & Adjectives   & 22          & .95 (20) & .95    & .68  & .32  & .73 (12) \\
6 & Comparatives & 31          & .87      & .56    & .48  & .45  &  -       \\
8 & Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
9 & Attitudes    & 13          & .92      & .85    & .77  & .46  & .92 (9)  \\ \hline
  & Total        & 262         & .89      & .83    & .69  & .50  & .85  \\
  &              &             & (259)    & (174)  & (174)& (174)& (89)
  \end{tabular}
  \caption{Overview of the accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of examples shown in the
    third column, the actual number of examples attempted is shown in
    parentheses.  ``FraCoq'' refers to the work of
    \citet{bernardy_type_2017}. ``MINE" refers to the approach of
    \citet{Mineshima:2015}, ``NUT" to the CCG system that utilizes the
    first-order automated theorem prover \textit{nutcracker} described
    by \citet{bos:2008}, and ``Langpro" to the system presented by
    \citet{Abzianidze:2015}.  }
\end{table}

The above result reveals a considerable improvement over earlier
approaches in terms of coverage, with three more sections covered over
previous state of the art. Additionally, our system is generally best
in terms of accuracy overall. In particular, section 6 sees a large
improvement in accuracy, which we attribute to our new analysis of
comparatives, enabled by our dynamic semantics.

\todo{error analysis}

>>>>

\usepackage{stmaryrd}
\usepackage{enumitem}
\usepackage{newunicodechar}
\input{PaperTools/latex/newunicodedefs}

@@

\newlist{lingex}{enumerate}{3}
\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}

@@

The dynamic semantics follows a monadic structure, as pioneered by
\citet{Shan:2002}. There are two kinds of effects carried by the
monad.  The first one comprises a series of updates and queries of
stateful elements.  There is one piece of updateable state for every
element which can be referred to by anaphoric expressions. These can
be the usual ones (like NPs), but also less usual ones (like 2-place
verbs, or a quantity --- which we illustrate below). The other kind of
effects is \emph{non-determinism}. We use non-determinism to model the
property that linguistic expressions can have several
interpretations. The monadic structure allows to locally express that
a given expression has several meanings; the monadic bind ensures that
all combinations of meanings are considered at the top-level,
combinatorially. This dynamic semantics allows us to model many
phenomena in a precise way.

\paragraph{anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples such as the following:
\item[FraCaS 122]
P1	Every committee has a chairman.\\
P2	He is appointed its members.\\
H 	Every committee has a chairman appointed by members of the committee. 
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase. Such
ellptic expressions are handled in the same way as anaphora.
\begin{lingex}
\item[FraCas 173]
  P1	Bill spoke to everyone that John did [elliptic V2].\\
  P2	John spoke to Mary.\\
  Q 	Did Bill speak to Mary?\\
  H 	Bill spoke to Mary.
\end{lingex}
For example, in (FraCaS 173, P1), the argument of ``did'' is
explicitly marked as an elliptic V2, which we resolve to ``speak'' in
that context.
\paragraph{definites} A naive way to handle definites is an
existential (or sigma). However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

Instead, our take is that definites should be treated as an anaphoric
expression with an implicit referrent. That is, if the referent is not
found in the discourse, then it will be forcibly introduced, using an
existential type, \emph{at the top-level} of the expression. To be
able to do this, we record all definites without referent, using
another portion of the environment (using a monadic effect). For the
above example, we obtain the desired interpretation:
$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions. Finally, ``more'' is intepreted as ``at least
<elliptic quantity>+1''. This treatment, which we illustrated here on
an example, is systematic in our implementation.

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
\todo{reference to this view in the litterature}.  Additionally,
adjectival phrases, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable. Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

\paragraph{Adverbs}
Another point, of minor theoretical importance but major practical
one, is our handling of adverbial phrases. We interpret adverbs (and
in general all adverbial and prepositional phrases) as VP-modifiers:
$Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
to verb-phrases heavily complicates the Coq proofs, because such
phrases can contain quantifiers. Therefore, we instead move the
adverbs, so that they apply to (atomic) verbs only. Proofs can then be
simplify accordingly. \todo{theoretical justification in the
  literature?}

@@

We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
total of 259 examples. We excluded only section 7, ``temporal reference''. The reason for
doing so is that, in our view, it contains too many examples which
require \textit{ad-hoc} treatment, and thus makes little sense to
include without complementing it with a more thorough suite which
captures a more complete landscape of the phenomena that section 7
touches.

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by Bill MacCartney\todo{cite} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labelling (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of the following cases, which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.

@@

Our system classifies a case as YES if a proof can be constructed from
the premises to the hypothesis, NO if a proof of the negated
hypothesis can be constructed and UNK otherwise. Because we work with
a non-decidable logic, one cannot \emph{in general} conclude
decisively that no proof exists. Thus, we consider here that no proof
exists if it cannot be constructed with reasonable effort. In
particular, we test at the minimum that the automatic proof search
built in Coq does not succeed before classifying a problem as UNK.

@@

\providecommand\forcecenter{\multicolumn{1}{c}}
\begin{table}
  \centering
\begin{tabular}{rlrrrrrr}
  & Section      & \# examples & Ours     & FraCoq & MINE & Nut  & Langpro  \\ \hline
1 & Quantifiers  & 75          & .96 (74) & .96    & .77  & .53  & .93 (44) \\
2 & Plurals      & 33          & .82      & .76    & .67  & .52  & .73 (24) \\
3 & Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
4 & Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
5 & Adjectives   & 22          & .95 (20) & .95    & .68  & .32  & .73 (12) \\
6 & Comparatives & 31          & .87      & .56    & .48  & .45  &  -       \\
7 & Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
8 & Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
9 & Attitudes    & 13          & .92      & .85    & .77  & .46  & .92 (9)  \\ \hline
  & Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
  &              &             & (259)    & (174)  & (174)& (174)& (89)
  \end{tabular}
  \caption{Overview of the accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of examples shown in the
    third column, the actual number of examples attempted is shown in
    parentheses.  ``FraCoq'' refers to the work of
    \citet{bernardy_type_2017}. ``MINE" refers to the approach of
    \citet{Mineshima:2015}, ``NUT" to the CCG system that utilizes the
    first-order automated theorem prover \textit{nutcracker} described
    by \citet{bos:2008}, and ``Langpro" to the system presented by
    \citet{Abzianidze:2015}. A dash indicates that no attempt was made
    for a given section. }
  \label{tbl:results}
\end{table}

Table \ref{tbl:results} shows a considerable improvement over earlier
approaches in terms of coverage, with three more sections covered over
previous state of the art. We thus cover 259 out of 337 cases (77\%),
compared to at most 174 cases (52\%) in previous work. Additionally,
our system performs generally the best in terms of accuracy. In
particular, section 6 sees a large improvement in accuracy, which we
attribute to our new analysis of comparatives, enabled by our dynamic
semantics.

\paragraph{error analysis}
Our system fails to classify correctly 28 cases out of 259.  We give
here a summary of the missing features which are responsible for the
biggest number of failures. The one biggest source of error is an
incomplete handling of group readings. (FraCaS
013,014,046,084,111,124,126,127,137,171,172,191,193,195,243,250,333,346). These
are cases where a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases are where a definite plural does
not have a universal reading (091,094,095) --- these require an
analysis at the pragmatic level, which our system lacks. Additionally,
neither measure phrases (242) nor attributive comparatives (244,245)
are handled.
